|

### 2022-05-16 Memo/memo.Rmd
update memo and appendix

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author: Response to Reviewers
date: \today
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: single
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: no
    keep_tex: no
    toc: no
    toc_depth: 3
    template: memo-template.latex
    md_extensions: +raw_attribute
---

@@

In response to comments from Reviewer 1, we have revised our Methods section to
clarify our selection of independent variables, including the use of dummy
variables to improve the accuracy of our estimates, and to provide further
information about our online survey. Additionally, we have included a new
section in the Supplemental Materials to explain how we conducted our survey
and how it contributes to our results. We believe these additions also give the
paper a better theoretical frame to help readers sort through the (many)
results we present.

@@

We thank Reviewer 1 for these two relevant suggestions. We now cite Gilligan
and Matsusaka's 2001 article three times in our paper. We mention it twice in
our Introduction, first on page 2: ``\textit{On the one hand, enlarged
legislatures may reduce political particularism and increase overall government
spending, which can lead to more public goods (Gerring and Veenendaal 2020;
Gilligan and Matsusaka 2001; Weingast et al. 1981)}'', and then on page 4:
``\textit{The corollary of the "law of $1/n$" is that increasing legislature
size promotes excessive public spending, regardless of the partisan makeup of
the legislature (Gilligan and Matsusaka 2001)}''. We also cite it in our Theory
section (page 7): ``\textit{Yet, Gilligan and Matsusaka (2001) find that in the
US legislature size is associated primarily with higher spending on
geographically concentrated benefits, such as education and highways}''.

We cite Gerring and Veenendaal's book four times. We mention it twice on page
2: ``\textit{Legislature size is known to have a significant effect on
collective action strategies, the bargaining power between state branches,
government spending, and representation (Crain 1979; Freire et al. 2021;
Gerring and Veenendaal 2020; Primo and Snyder Jr. 2008; Rogers 2002; Weingast
et al. 1981). However, how legislature size impacts social welfare remains less
understood. On the one hand, enlarged legislatures may reduce political
particularism and increase overall government spending, which can lead to more
public goods (Gerring and Veenendaal 2020; Gilligan and Matsusaka 2001;
Weingast et al. 1981)}''. We mention the book again on page 4:
``\textit{Following the seminal contribution by Weingast et al. (1981), the
"law of $1/n$" posits that larger legislatures increase the legislative pressure
for providing targeted public goods to local supporters at the expense of the
entire polity (Freire et al. 2021). This question has become more relevant
since legislature size has increased consistently over time (Gerring and
Veenendaal 2020)}''. Finally, we also cite it on page 6: ``\textit{Larger
legislatures improve representation as both the election thresholds and the
ratio of population-per-politician decrease, which facilitate the electoral
success of ethnic and political minorities (Allen and Stoll 2014; Gerring and
Veenendaal 2020)}.''

@@

We thank Reviewer 1 for the opportunity to clarify the selection of these two
sets of variables for our paper. We explain our decision on page 17:
"\textit{Our focus is on the impact of legislature size on education and health
care, which are the two most important public services the municipal government
should provide to citizens. According to Brazilian Constitution, municipalities
must spend 25\% of their budget on education and 15\% on health care.
Consequently, the Ministry of Health and the Ministry of Education provide
readily available, fine-grained data on health care and education at the
municipal level, which we leverage in this study}."

@@

We agree with Reviewer 1 that we needed to provide more information on the
online survey. Therefore, we have detailed our procedures in Section A.14 of
the Supplementary Materials (pages 22-23), and we also reference the survey on
pages 17 and 22 (footnote 13) in the main article.

@@

We thank Reviewer 1 for the opportunity to clarify the selection of control variables in our empirical analysis. The inclusion of controls in our model has two purposes. The first one is to improve the consistency of the estimates. In section A.5 of the Appendix, we show by simulation that adding controls enhances the consistency of the point estimates. This is true for several distributions of the running variables and many assumptions regarding the shape of the treatment effects. Section 5 of the Main Paper demonstrates that it improves consistency significantly in both the main estimates and the placebo ones. Table 1 in Section 5 of the Main Paper (reproduced below), shows that without controls, we would overestimate the treatment effects (finding an increase of 1.63 legislators instead of a sharp increase of one legislator), and underestimate the placebo treatment effects (finding a significant decrease of 0.31 legislators where there was supposed to be no effects in legislature size). This substantial improvement is because having control variables makes the subpopulations in each threshold more comparable across all the thresholds, avoiding comparing cases to the left (right) of one given threshold with points in the right (left) of a different threshold.

@@

The second reason, per \citet{calonico2019regression} results, is that adding control variables improves the efficiency of our estimates. We add five controls to the main specifications. We added the population in 2003 because it conditions the multiple assignments. Adding the GDP per capita ensures that the gains detected are not resulting from the hot job market experienced during the period. We add the number of seats in 2000 to ensure that the results come from the current legislature size decision and not from their past seat allocation. We add the year to control for likely seasonal effects. Finally, we added a dummy for Northeast states. The Northeast region received considerable attention from the Federal Government in the 2000s. According to \citet{de2015politica}, the Lula government in 2003 recreated the Superintendency for the Development of the Northeast (SUDENE), investing BRL 84 Billion from 2003 to 2010. This amounts to around 70% of all targeted regional investments. One example of these investments is the number of Federal Educational Institutions: the Federal Government created about 100 university campuses in the region. Moreover, infant mortality in the Northeast region decreased from 22.4 deaths per 1000 infants born in 2005 to 19.6 deaths per 1000 infants born in 2008. Education quality improved from a 2.69 score in 2005 to a 3.14 score in 2007 (0 to 10 quality of education score). For this reason, adding a dummy for Northeast states is essential for ensuring the consistency and the efficiency of our estimates.

To reflect these pieces of information, we rewrote paragraph five in Section 5 of the Main Text to:

``\textit{(...) In the dataset, there are twelve discontinuities. At
each threshold, there is a sharp increase of one councilor, from 9 to 21. When
running an RDD on these discontinuities, a consistent estimator should fit the
exact change of one councilor. However, if we pool all the discontinuities
together with no correction, we find an increase in council size of 1.63
councilors (see the first estimate in Table 1). This
imbalance occurs because when we combine all the discontinuities, the model
implies that municipalities below the 47,619 cutoffs (9 to 10 councilors) are
comparable to changes in towns right above the 571,428 cutoffs (20 to 21
council members). That is incorrect because the comparison is inconsistent in
council size, and these municipalities diverge in many other aspects.}

\textit{To avoid this problem, in section A.5 of the Supplemental Materials, we show by
simulation that adding controls, especially the variables responsible for the
multiple threshold assignments (in our case, the population in 2003), improves
the consistency and efficiency of the estimates. Relying on this fact, we add
five controls to the estimates: population in 2003; GDP per capita; the number
of seats in 2000; year; and a dummy for Northeast municipalities. The
reason for the first variable is the multiple threshold assignments. We add GDP
per capita because wealthier cities tend to have better public services. The
number of seats in 2000 intends to control that some municipalities could
experience a change provoked by the previous council size, confounding our
estimates. The Northeast dummy was added to control for the fact that the Lula
administration heavily invested in the region from 2003 until 2010, improving 
several development indicators \citep{de2015politica}. The years were added 
to improve efficiency and control for seasonal effects. Moreover, adding controls improves efficiency in RDD \citep{calonico2019regression} and also enhances the consistency in the multiple thresholds RDD.}''

@@

We agree with R2 that testing the proposed mechanism is an extremely exciting
project for a follow-up article. In the Discussion section (page 29), we write:
"\textit{Second, future research could evaluate whether additional legislators
help mayors gain access to federal or state transfers in countries where
resources are centrally provided, such as Japan. As suggested by Catalinac et
al. (2020), in these situations the national government can create a tournament
between municipalities and provide resources to those which offer more
electoral support. Scholars might want to analyze whether this tournament
structure interacts with the mechanisms we suggest in this paper}."

@@

We collapse two concerns from Reviewer 3 since they both speak to core topics
in our paper. In Table \textbf{XX} of the Supplementary Materials (page
\textbf{XX}), we show that an additional legislator usually does not lead to
significant increases in public expenditure. Therefore, our results speak more
about the allocation of resources than about government spending. In our
section titled "Background on the Brazilian Case" (page 15), we clarify why
increasing the city council size often results in a pro-mayor legislator. We
write: "\textit{Brazil also adopts a unique formula to allocate seats, which
combines the Hare quota and the D’Hondt method and strongly favors larger
parties (Nicolau 2015). This poses a major obstacle for small parties to reach
the electoral quorum, so they often join party coalitions to improve their
chances of winning legislative seats (Ames 2001, p. 68). In Brazil, mayors are
usually members of the largest coalitions (Poulsen and Varjao 2018), thus there
is an incentive for candidates to side with the mayor before and after the
election}." We mention that information again on page 22, footnote 12:
"\textit{The effect of city council size on the number of mayor-aligned
politicians stems from the rounding rule adopted by the proportional
representation system in Brazil. The rounding rule favors the coalitions with
more votes. Most of the time, the coalition of the elected mayor is the one
that receives more votes.}".

We did not find evidence that electing an additional legislator improves
descriptive representation in the city council. For instance, we see no
significant increase in the number of more female or non-white legislators, as
shown in Table 4 (page 27). Finally, we show in Section \textbf{AAA} of the
Supplementary Materials (page \textbf{XX}) that the mayors' parties are driving
our results.

@@

We entirely agree with Reviewer 3 that we need to discuss the scope conditions
of our findings in more detail. We acknowledge that the external validity of
our findings might be conditional on the contextual characteristics of our
study. We have rewritten two paragraphs in the Discussion (page 29) to highligh
the implications and limits of our findings. We write: "\textit{Our findings
have implications that extend well beyond Brazil. A general form of our
argument is that it is easy for a politician to negotiate with members of her
in-group (Abbink and Harris 2019; Alt et al. 2018). Specifically, several
countries have separate executive and legislative powers, such as the federal
level of presidential systems, state governments, and local city councils. As
these political bodies are influenced by political allegiances, the dynamics we
expose here may be generalizable. Depending on legislative support, changing
the legislature size could increase the chance of electing politicians that are
not aligned with the executive, hence decreasing welfare. Despite the
particular effects on political support and welfare, the logic described in
this paper would remain relevant for empirically accessing the effect of
enlarged legislatures on welfare. Our findings may also help scholars
understand why other countries in Latin America and Asia can generate effective
governance despite having strong executives and large coalitions (Pereira and
Melo 2012)}".

In the next paragraph (page 29), we indicate some of the limitations of our
findings and encourage scholars to conduct further research on the topic. We
write: "\textit{This paper opens several questions for future analysis. [...]
Third, scholars might want to investigate whether our findings remain valid
under different conditions. It is possible that the mechanisms we present here
are non-linear, with additional legislators having a positive or negative
impact on welfare depending on certain electoral characteristics. Finally, it
remains unclear how other legislative features affect citizen well-being. For
example, city councils may vary in size, monetary compensation for councilors,
committee structure, and internal power structure. Understanding how
legislature size interacts with other features would improve our knowledge
about the impact of local and national legislatures impact public services.
These questions are crucial for institution design and welfare in developing
democracies.}"

@@

>>>>

---
title:  |
        | Legislature Size and Welfare: Evidence from Brazil
author: Response to Reviewers
date: \today
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: single
papersize: a4paper
bibliography: ../article/references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: no
    keep_tex: no
    toc: no
    toc_depth: 3
    template: memo-template.latex
    md_extensions: +raw_attribute
---

@@

In response to comments from Reviewer 1, we have revised our Methods section to
clarify our selection of independent variables, including the use of covariates
to improve the efficiency and consistency of our estimates. Additionally, we
have included a new section in the Supplemental Materials to explain how we
conducted our survey and how it contributes to our results. We believe these
additions also give the paper a better theoretical frame to help readers sort
through the results we present.

@@

We thank Reviewer 1 for these two relevant suggestions. We now cite Gilligan and
Matsusaka's 2001 article three times in our paper. We mention it twice in our
Introduction, first on page 2: ``\textit{On the one hand, enlarged legislatures
may reduce political particularism and increase overall government spending,
which can lead to more public goods  \citep{gerring2020population,
GilliganMatsusaka2001, Weingastetal1981}}'', and then in the end of page 4:
``\textit{The corollary of the "law of $1/n$" is that increasing legislature
size promotes excessive public spending, regardless of the partisan makeup of
the legislature \citep{GilliganMatsusaka2001}}''. We also cite it in our Theory
section (page 7): ``\textit{Yet, \citet{GilliganMatsusaka2001} find that in the
US legislature size is associated primarily with higher spending on
geographically concentrated benefits, such as education and highways}''.

We cite Gerring and Veenendaal's book four times. We mention it twice on page 2:
``\textit{Legislature size is known to have a significant effect on collective
action strategies, the bargaining power between state branches, government
spending, and representation \citep{crain1979cost, freire2021legislature,
gerring2020population, PrimoSnyder2008, rogers2002free, Weingastetal1981}.
However, how legislature size impacts social welfare remains less understood. On
the one hand, enlarged legislatures may reduce political particularism and
increase overall government spending, which can lead to more public goods
\citep{gerring2020population, GilliganMatsusaka2001, Weingastetal1981}}''. We
mention the book again on page 4: ``\textit{Following the seminal contribution
by \citet{Weingastetal1981}, the "law of $1/n$" posits that larger legislatures
increase the legislative pressure for providing targeted public goods to local
supporters at the expense of the entire polity \citep{freire2021legislature}. This
question has become more relevant since legislature size has increased
consistently over time \citep{gerring2020population}}''. Finally, we also cite
it on page 6: ``\textit{Larger legislatures improve representation as both the
election thresholds and the ratio of population-per-politician decrease, which
facilitate the electoral success of ethnic and political minorities
\citep{allenrepresentation, gerring2020population}}.''

@@

We thank Reviewer 1 for the opportunity to clarify the selection of these two
sets of variables for our paper. We included an explanation for our choice on
page 17: "\textit{Our focus is on the impact of legislature size on education
and health care, which are the two most important public services the municipal
government should provide to citizens. According to the Brazilian Constitution,
municipalities must spend 25\% of their budget on education and 15\% on health
care. Consequently, the Ministry of Health and the Ministry of Education provide
readily available, fine-grained data on health care and education at the
municipal level, which we leverage in this study}."

@@

We agree with Reviewer 1 that we needed to provide more information on the
online survey. Therefore, we have detailed our procedures in Section A.14 of the
Supplementary Materials (pages 22-23), and we also reference the survey on pages
17 and 22 (footnote 13) in the main article. We rewrote paragraph 2 of Section
6.2 to clarify the importance of the survey results in our paper. It reads as
follows: ``\textit{We assume that city councilors want to extract rents and to
get reelected. The rents come from continuing in the repeated game, and the
partisanship discounts that benefit. Figure 2 shows that 65.9 percent of
councilors mentioned that mayors use job appointments to consolidate their
coalition support. Moreover, 59.0 percent of the surveyed councilors affirmed
that mayors use personal favors, such as shifting resources to places where the
councilors have electoral support, to solidify their governing coalition. This
provides further evidence that mayors do offer city councilors the benefits they
demand. In this regard, granting favors works as rent transfers for councilors.
Job appointments are also convenient for the mayors because patronage in Brazil
has positive results in terms of welfare \citep{toral2019benefits}.}''

@@

We thank Reviewer 1 for the opportunity to clarify the selection of control
variables in our empirical analysis. The inclusion of controls in our model has
two purposes. The first one is to improve the consistency of the estimates. In
Section A.5 of the Appendix, we use simulation to show that the inclusion of
covariates enhances the consistency of our point estimates. The simulation
indicates improved consistency for several distributions of the running
variables and different treatment effects. Section 5 of the manuscript
demonstrates that it improves consistency in both the main estimates and the
placebo ones. Table 1 (page 20, reproduced below) shows that without controls,
we would overestimate the treatment effects (finding an increase of 1.63
legislators instead of an increase of exactly one legislator), and underestimate
the placebo treatment effects (finding a significant decrease of 0.31
legislators where there was supposed to be no effects in legislature size). This
substantial improvement is because having control variables makes the
subpopulations in each threshold more comparable across all the thresholds,
avoiding comparing cases to the left (right) of one given threshold with points
in the right (left) of a different threshold.

@@

The second reason is that adding control variables improves the efficiency of
our estimates \citep{calonico2019regression}. We add five controls to the main
specifications. We include the 2003 population because it conditions the
multiple threshold assignments. We add GDP per capita to control for Brazil's
economic expansion during the 2000s. We also add the number of seats in 2000 to
ensure that our results come from the current legislature size and not from past
seat allocations. Year was included in the models to control for seasonal
effects. Finally, we add a dummy for Northeast states. The Northeast region
received considerable attention from the Federal Government in the 2000s.
According to \citet{de2015politica}, in 2003 the Lula administration recreated
the Superintendency for the Development of the Northeast (SUDENE), investing BRL
84 Billion from 2003 to 2010. This amounts to around 70% of all targeted
regional investments. One example of these investments is that the Federal
Government created about 100 new universities in the Northeast region. Moreover,
infant mortality in the region decreased from 22.4 deaths per 1000 infants born
in 2005 to 19.6 deaths per 1000 infants born in 2008 due to larger federal
investments. In turn, education quality improved from 2.69 in 2005 to a 3.14
score in 2007 (0 to 10 quality of education score). For this reason, adding a
dummy for Northeast states is essential for ensuring the consistency and the
efficiency of our estimates.

To reflect these pieces of information, we rewrote paragraph five in Section 5
of the manuscript to (page 19):

``[...] \textit{we add five controls to the estimates: population in 2003; GDP
per capita; the number of seats in 2000; year; and a dummy for Northeast
municipalities. The reason for the first variable is the multiple threshold
assignments. We add GDP per capita because wealthier cities tend to have better
public services. The number of seats in 2000 intends to control that some
municipalities could experience a change provoked by the previous council size,
confounding our estimates. The Northeast dummy was added to control for the fact
that the Lula administration heavily invested in the region from 2003 until
2010, improving several development indicators \citep{de2015politica}. The years
were added to improve efficiency and control for seasonal effects. Moreover,
adding controls improves efficiency in RDD \citep{calonico2019regression} and
also enhances the consistency in the multiple thresholds RDD.}''

@@

We agree with Reviewer 2 that testing the proposed mechanism is an exciting
project for a follow-up article. In the Discussion section (page 29), we write:
"\textit{Second, future research could evaluate whether additional legislators
help mayors gain access to federal or state transfers in countries where
resources are centrally provided, such as Japan. As suggested by
\citet{catalinac2020tournament}, in these situations the national government can
create a tournament between municipalities and provide resources to those which
offer more electoral support. Scholars might want to analyze whether this
tournament structure interacts with the mechanisms we suggest in this paper}."

@@

We collapse two concerns from Reviewer 3 since they both speak to core topics in
our paper. In Table 6 of the Supplementary Materials (page 21), we show that an
additional legislator usually does not lead to significant increases in public
expenditure. Therefore, our results speak more about the allocation of resources
than about government spending. In our section titled "Background on the
Brazilian Case" (page 14), we clarify why increasing the city council size often
results in a pro-mayor legislator. We write (page 15): "\textit{Brazil also
adopts a unique formula to allocate seats, which combines the Hare quota and the
D’Hondt method and strongly favors larger parties \citep{nicolau2015sistemas}.
This poses a major obstacle for small parties to reach the electoral quorum, so
they often join party coalitions to improve their chances of winning legislative
seats \citep[68]{Ames2001}. In Brazil, mayors are usually members of the largest
coalitions \citep{poulsen2018political}, thus there is an incentive for
candidates to side with the mayor before and after the election}." We mention
that information again on page 22, footnote 12: "\textit{The effect of city
council size on the number of mayor-aligned politicians stems from the rounding
rule adopted by the proportional representation system in Brazil. The rounding
rule favors the coalitions with more votes. Most of the time, the coalition of
the elected mayor is the one that receives more votes.}".

We did not find evidence that electing an additional legislator improves
descriptive representation in the city council. For instance, we see no
significant increase in the number of female or non-white legislators, as shown
in Table 4 (page 27). Finally, we show in the table below that the mayors'
parties are not driving our results. This result is also available on our GitHub
repository. We ommitted it from the Supplementary Materials to conserve space.

\begin{table}[!htbp] \centering 
\caption{Councilors from Major Parties} 
\label{appcouncilparty} 
\resizebox{\textwidth}{!}{
  \begin{tabular}{@{\extracolsep{\fill}}l c c c c} 
  \\[-1.8ex]\hline \\[-1.8ex] 
  & \multicolumn{1}{c}{\textbf{Councilors from PT}} & \multicolumn{1}{c}{\textbf{Councilors from PSDB}} & \multicolumn{1}{c}{\textbf{Councilors from PMDB}} & \multicolumn{1}{c}{\textbf{Councilors from PFL (DEM)}} \\ 
  \hline \\[-1.8ex] 
  LATE & -0.01 & -0.05 & -0.01 & -0.01 \\ 
  & (0.02) & (0.05) & (0.03) & (0.02) \\ 
  N Left & 5179 & 5179 & 5179 & 5179 \\ 
  N Right & 343 & 343 & 343 & 343 \\ 
  Eff N Left & 375 & 152 & 267 & 231 \\ 
  Eff N Right & 212 & 121 & 172 & 159 \\ 
  BW Loc Poly & 13.99 & 6.8 & 10.58 & 9.63 \\ 
  BW Bias & 22.64 & 11.373 & 16.803 & 15.647 \\ 
  \hline \\[-1.8ex] 
  \multicolumn{5}{l}{$^{***}$p $<$ .01; $^{**}$p $<$ .05; $^{*}$p $<$ .1} \\ 
  \end{tabular} }
\end{table}

@@

We entirely agree with Reviewer 3 that we need to discuss the scope conditions
of our findings in more detail. We acknowledge that the external validity of
our results might be conditional on the contextual characteristics of our
study. We have rewritten two paragraphs in the Discussion (page 29) to highligh
the implication of our results. We write: "\textit{Our findings
have implications that extend well beyond Brazil. A general form of our
argument is that it is easy for a politician to negotiate with members of her
in-group \citep{abbink2019, g9030057}. Specifically, several
countries have separate executive and legislative powers, such as the federal
level of presidential systems, state governments, and local city councils. As
these institutions are influenced by political allegiances, the dynamics we
expose here may be generalizable. Depending on legislative support, changing
the legislature size could increase the chance of electing politicians that are
not aligned with the executive, hence decreasing welfare. Despite the
particular effects on political support and welfare, the logic described in
this paper would remain relevant for empirically accessing the effect of
enlarged legislatures on welfare. Our findings may also help scholars
understand why other countries in Latin America and Asia can generate effective
governance despite having strong executives and large coalitions
\citep{pereira2012surprising}.}".

In the next paragraph (page 29), we indicate some of the limitations of our
findings and encourage scholars to conduct further research on the topic. We
write: "\textit{This paper opens several questions for future analysis. [...]
Third, scholars might want to investigate whether our findings remain valid
under different conditions. It is possible that the mechanisms we present here
are non-linear, with additional legislators having a positive or negative
impact on welfare depending on certain electoral characteristics. Finally, it
remains unclear how other legislative features affect citizen well-being. For
example, city councils may vary in size, monetary compensation for councilors,
committee structure, and internal power structure. Understanding how
legislature size interacts with other features would improve our knowledge
about how local and national legislatures impact public services.
These questions are crucial for institution design and welfare in developing
democracies.}"

@@

\newpage

### 2022-05-16 appendix/appendix.Rmd
update memo and appendix

>>>>

The inclusion of controls in our model has two purposes. The first one is to
improve the consistency of the estimates. The second reason is that adding
control variables improves the efficiency of our estimates
\citep{calonico2019regression}. We add five controls to the main
specifications. We include the 2003 population because it conditions the
multiple threshold assignments. We add GDP per capita to control for Brazil's
economic expansion during the 2000s. We also add the number of seats in 2000 to
ensure that our results come from the current legislature size and not from
past seat allocations. Year was included in the models to control for seasonal
effects. Finally, we add a dummy for Northeast states. The Northeast region
received considerable attention from the Federal Government in the 2000s.
According to \citet{de2015politica}, in 2003 the Lula administration recreated
the Superintendency for the Development of the Northeast (SUDENE), investing
BRL 84 Billion from 2003 to 2010. This amounts to around 70% of all targeted
regional investments. One example of these investments is that the Federal
Government created about 100 new universities in the Northeast region.
Moreover, infant mortality in the region decreased from 22.4 deaths per 1000
infants born in 2005 to 19.6 deaths per 1000 infants born in 2008 due to larger
federal investments. In turn, education quality improved from 2.69 in 2005 to a
3.14 score in 2007 (0 to 10 quality of education score). For this reason,
adding a dummy for Northeast states is essential for ensuring the consistency
and the efficiency of our estimates.

### 2022-05-16 article/article.Rmd
update memo and appendix

The last set of variables quantifies social welfare. Our focus is on the impact
of legislature size on education and health care, which are the two most
important public services the municipal government should provide to citizens.
According to Brazilian Constitution, municipalities must spend 25\% of their
budget on education and 15\% on health care. Consequently, the Ministry of
Health and the Ministry of Education provide readily available, fine-grained
data on health care and education at the municipal level, which we leverage in
this study. We measure access to education with the average enrollment in
primary (K--4) municipal public schools. We assess education quality with the
Education Development Index (Índice de Desenvolvimento da Educação
Básica--IDEB), which the Ministry of Education has issued biannually since
2007. The index is a weighted average of student scores in Portuguese and
Mathematics (0--10), multiplied by the harmonic mean of public school
promotion rates in a given year (0--100). To test the effect of the 2004
ruling on health care, we collect data on infant mortality per thousand
children born and on postneonatal infant mortality, which is the death of
infants who survived more than 28 days, but died before reaching one year
of age. The source of all health care variables is the Brazilian Ministry
of Health Data Center (DataSUS).

@@

We assume that city councilors want to extract rents and to get reelected.
The rents come from continuing in the repeated game, and the partisanship
discounts that benefit. Figure \ref{survelec} shows that 65.9 percent of
councilors mentioned that mayors use job appointments to consolidate their
coalition support. Moreover, 59.0 percent of the surveyed councilors affirmed
that mayors use personal favors, such as shifting resources to places where the
councilors have electoral support, to solidify their governing coalition.
Granting favors works as rent transfers for councilors. Job appointments are
also convenient for the mayors because patronage in Brazil has positive results
in terms of welfare \citep{toral2019benefits}.

@@

Our findings have implications that extend well beyond Brazil. A general form
of our argument is that it is easy for a politician to negotiate with members
of her in-group \citep{abbink2019, g9030057}. Specifically, several countries
have separate executive and legislative powers, such as the federal level of
presidential systems, state governments, and local city councils. As these
political bodies are influenced by political allegiances, the dynamics we
expose here may be generalizable. Depending on legislative support, changing
the legislature size could increase the chance of electing politicians that are
not aligned with the executive, hence decreasing welfare. Despite the
particular effects on political support and welfare, the logic described in
this paper would remain relevant for empirically accessing the effect of
enlarged legislatures on welfare. Our findings may also help scholars
understand why other countries in Latin America and Asia can generate effective
governance despite having strong executives and large coalitions
\citep{pereira2012surprising}.

This paper opens several questions for future analysis. First, if the mechanism
we suggest is indeed at work, one may ask what should be the city council's
checking prerogatives, as they significantly reduce the provision of public
services. The evidence so far is inconclusive, and \citet{poulsen2018political}
show that checks for the mayors may also improve service provision. Second,
future research could evaluate whether additional legislators help mayors gain
access to federal or state transfers in countries where resources are centrally
provided, such as Japan. As suggested by \citet{catalinac2020tournament}, in
these situations the national government can create a tournament between
municipalities and provide resources to those which offer more electoral
support. Scholars might want to analyze whether this tournament structure
interacts with the mechanisms we suggest in this paper. Third, scholars might
want to investigate whether our findings remain valid under different
conditions. It is possible that the mechanisms we present here are non-linear,
with additional legislators having a positive or negative impact on welfare
depending on certain electoral characteristics. Finally, it remains unclear how
other legislative features affect citizen well-being. For example, city
councils may vary in size, monetary compensation for councilors, committee
structure, and internal power structure. Understanding how legislature size
interacts with other features would improve our knowledge about the impact of
local and national legislatures impact public services. These questions are
crucial for institution design and welfare in developing democracies.

>>>>

The last set of variables quantifies social welfare. Our focus is on the impact
of legislature size on education and health care, which are the two most
important public services the municipal government should provide to citizens.
According to the Brazilian Constitution, municipalities must spend 25\% of their
budget on education and 15\% on health care. Consequently, the Ministry of
Health and the Ministry of Education provide readily available, fine-grained
data on health care and education at the municipal level, which we leverage in
this study. We measure access to education with the average enrollment in
primary (K--4) municipal public schools. We assess education quality with the
Education Development Index (Índice de Desenvolvimento da Educação
Básica--IDEB), which the Ministry of Education has issued biannually since
2007. The index is a weighted average of student scores in Portuguese and
Mathematics (0--10), multiplied by the harmonic mean of public school
promotion rates in a given year (0--100). To test the effect of the 2004
ruling on health care, we collect data on infant mortality per thousand
children born and on postneonatal infant mortality, which is the death of
infants who survived more than 28 days, but died before reaching one year
of age. The source of all health care variables is the Brazilian Ministry
of Health Data Center (DataSUS).

@@

We assume that city councilors want to extract rents and to get reelected. The
rents come from continuing in the repeated game, and the partisanship discounts
that benefit. Figure \ref{survelec} shows that 65.9 percent of councilors
mentioned that mayors use job appointments to consolidate their coalition
support. Moreover, 59.0 percent of the surveyed councilors affirmed that mayors
use personal favors, such as shifting resources to places where the councilors
have electoral support, to solidify their governing coalition. This provides
further evidence that mayors do offer city councilors the benefits they demand.
In this regard, granting favors works as rent transfers for councilors. Job
appointments are also convenient for the mayors because patronage in Brazil has
positive results in terms of welfare \citep{toral2019benefits}.

@@

Our findings have implications that extend well beyond Brazil. A general form
of our argument is that it is easy for a politician to negotiate with members
of her in-group \citep{abbink2019, g9030057}. Specifically, several countries
have separate executive and legislative powers, such as the federal level of
presidential systems, state governments, and local city councils. As these
institutions are influenced by political allegiances, the dynamics we
expose here may be generalizable. Depending on legislative support, changing
the legislature size could increase the chance of electing politicians that are
not aligned with the executive, hence decreasing welfare. Despite the
particular effects on political support and welfare, the logic described in
this paper would remain relevant for empirically accessing the effect of
enlarged legislatures on welfare. Our findings may also help scholars
understand why other countries in Latin America and Asia can generate effective
governance despite having strong executives and large coalitions
\citep{pereira2012surprising}.

This paper opens several questions for future analysis. First, if the mechanism
we suggest is indeed at work, one may ask what should be the city council's
checking prerogatives, as they significantly reduce the provision of public
services. The evidence so far is inconclusive, and \citet{poulsen2018political}
show that checks for the mayors may also improve service provision. Second,
future research could evaluate whether additional legislators help mayors gain
access to federal or state transfers in countries where resources are centrally
provided, such as Japan. As suggested by \citet{catalinac2020tournament}, in
these situations the national government can create a tournament between
municipalities and provide resources to those which offer more electoral
support. Scholars might want to analyze whether this tournament structure
interacts with the mechanisms we suggest in this paper. Third, scholars might
want to investigate whether our findings remain valid under different
conditions. It is possible that the mechanisms we present here are non-linear,
with additional legislators having a positive or negative impact on welfare
depending on certain electoral characteristics. Finally, it remains unclear how
other legislative features affect citizen well-being. For example, city
councils may vary in size, monetary compensation for councilors, committee
structure, and internal power structure. Understanding how legislature size
interacts with other features would improve our knowledge about how local and
national legislatures impact public services. These questions are crucial for
institution design and welfare in developing democracies.
